---
title: "Ancovy_Bay_HW"
output: html_document
date: "2022-10-31"
---

This has Jeewantha's edits to Amanda's workbook to get the economic output from the MSEs
The new code starts from line 250 or so

# Load packages
```{r}
pacman::p_load("tidyverse","Rpath")
```
# Try out Rsim
## Example to run simulation with Rpath
Based off the vignette: https://github.com/NOAA-EDAB/Rpath/blob/master/vignettes/Rsim.Rmd

To run the MSEs I used the balanced model that Sean set up:
```{r}
source(here::here("data-raw/Ancovy_Bay_balanced.R"))
```

Set up basic simulation using Rpath model and estimated parameters
```{r}
ABsim <- rsim.scenario(AB, AB.params, years = 1:100) # Convert parameters to rates
```

Double sealers effort after 25 years
```{r}
ABsim_sealer <- adjust.fishing(ABsim, parameter = "ForcedEffort", group = 'sealers', sim.year = 25:100, value = 2)
```

Run modified simulation scenario
```{r}
ABsim_sealerRun <- rsim.run(ABsim_sealer, method = "RK4", years = 1:100)
```
Plot results
```{r}
rsim.plot(ABsim_sealerRun)
```

## Double MzeroMort setting for seals
Try adjust.scenario function Double the parameter for seals?
```{r}
ABsim_MzeroMort <- adjust.scenario(ABsim, parameter = 'MzeroMort', group = "seals", value = 2)
```
Run modified simulation scenario
```{r}
ABsim_MzeroMortRun <- rsim.run(ABsim_MzeroMort, method = "RK4", years = 1:100)
```
Plot results
```{r}
rsim.plot(ABsim_MzeroMortRun)
```


## Double the cod natural mortality
Try the adjust.forcing function, double cod mortality beginning year 25
```{r}
ABsim_codM <- adjust.forcing(ABsim, parameter = 'ForcedMort', 
               group = "cod", 
               value = 2, # Double
               sim.year = 25:100, # Change impacts years 25-100
               sim.month = 0) # All months impacted 
               
```

Run modified simulation scenario
```{r}
ABsim_codMRun <- rsim.run(ABsim_codM, method = "RK4", years = 1:100)
```
Plot results
```{r}
rsim.plot(ABsim_codMRun)
```



# Set up a full MSE loop
1) Set up initial scenario (same as line 16 above)
```{r}
ABsim <- rsim.scenario(AB, AB.params, years = 1:100) # Convert parameters to rates
```

2) Initialize simulation for first 25 years and full simulation
```{r}
AB.init <- rsim.run(ABsim, method = "AB", years = 1:25)
AB.full <- copy(AB.init)
```
3) Set up HCR
```{r}
# Get reference point by turning off fishing
gear <- AB.params$model[Type==3, Group]
for(igear in 1:length(gear)){
  AB.base <- adjust.fishing(ABsim, parameter = "ForcedEffort", group=gear[igear], value = 0, sim.year = 0:100)
}
AB.b0 <- rsim.run(AB.base, method = "RK4", 1:25)
# Cod B0 value
cod <- extract.node(AB.b0, 'cod') 
cod.b0 <- max(cod$Biomass)
# Cod ref pt = 1/2 B0
cod.ref <- 0.5*cod.b0 # I think this value is too high so cod get fished to extinction at the cod.ref
cod.ref=0.2
# Actually make function to modify effort of the Rsim scenario object
bio.rule <- function(Rsim.scenario, Rsim.run, group, gear, ref.point, year){
  group.num <- which(Rsim.scenario$params$spname == group)
  gear.num <- which(Rsim.scenario$params$spname == gear) - Rsim.scenario$params$NUM_BIO
  current.effort <- Rsim.scenario$fishing$ForcedEffort[(year-1)*12+1, gear.num]
  if(Rsim.run$end_state$Biomass[group.num] > ref.point){ # If below ref pt increase effort
    Rsim.scenario <- adjust.fishing(Rsim.scenario, 'ForcedEffort', group=gear, sim.year = year+1, value = current.effort*1.05)
  }
  if(Rsim.run$end_state$Biomass[group.num] < ref.point){ # If above ref pt reduce effort
    Rsim.scenario <- adjust.fishing(Rsim.scenario, 'ForcedEffort', group = gear, sim.year = year + 1, value = current.effort*0.75)
  }
  
  return(Rsim.scenario)
}
```

4) Run full MSE loop
```{r}
for(iyr in 25:99){ # Simulate years 26 through 100
  # HCR
  ABsim <- bio.rule(Rsim.scenario = ABsim, Rsim.run = AB.full, group='cod', gear='trawlers', ref.point=cod.ref, year=iyr)
  # Step forward
  AB.full <- rsim.step(ABsim, AB.full, method = "AB", year.end = iyr+1)
}
```

Plot results
```{r}
# All nodes
rsim.plot(AB.full)
# Only cod node
cod <- extract.node(AB.full, 'cod')
plot(cod$Biomass, xlab = "Month", ylab = "Biomass")
abline(h=cod.ref) # add ref pt
```





# Scenarios setup for HW
## Seals decline and management stays the same
Initial scenario
```{r}
ABsim <- rsim.scenario(AB, AB.params, years = 1:25) # Convert parameters to rates
```
Increase seal mortality by 50%  starting in year 2
```{r}
ABsim_sealMort <- adjust.forcing(ABsim, parameter = "ForcedMort", group = 'seals', sim.year = 2:25, value = 1.5)
```
Run modified simulation scenario
```{r}
ABsim_sealMortRun <- rsim.run(ABsim_sealMort, method = "RK4", years = 1:25)
```
Plot results
```{r}
# Plot all
rsim.plot(ABsim_sealMortRun)
# Plot fish
rsim.plot(ABsim_sealMortRun, spname = c("cod", "whiting", "mackerel", "anchovy", "shrimp"))
# Plot only seals and cod
rsim.plot(ABsim_sealMortRun, spname = c("seals", "cod"))
species <- c('whales', 'seals', 'cod', 'whiting','mackerel', 'anchovy', 'shrimp', 'benthos', 'zooplankton', 'phytoplankton', 'detritus')
spBiomass <- NULL
for(isp in species){
  stock <- extract.node(ABsim_sealMortRun, isp)
  spBiomass <- cbind(spBiomass, stock$Biomass)
}
colnames(spBiomass) <- species
Month <- c(1:(12*100))
spBiomass <- cbind(spBiomass, Month)
spBiomass <- spBiomass %>% 
  as.data.frame() %>% 
  pivot_longer(cols = all_of(species), names_to = "Species", values_to = "Biomass")
spBiomass %>% # filter(Species == 'whales') %>%
  ggplot() +
  geom_line(aes(x= Month, y = Biomass, col=Species)) +
  facet_wrap(~ Species, nrow = 3)
cod <- extract.node(ABsim_sealMortRun, 'cod')
plot(cod$Biomass, xlab = "Month", ylab = "Biomass")
abline(h=cod.ref) # add ref pt
```


## Seals decline starting in year 2 years and sealers harvest cut in half five years later (year 7)
```{r}
ABsim_sealMort <- adjust.forcing(ABsim, parameter = "ForcedMort", group = 'seals', sim.year = 2:25, value = 1.5)
ABsim_sealMortFish <- adjust.fishing(ABsim_sealMort, parameter = "ForcedEffort", group = 'sealers', sim.year = 7:25, value = 0.5)
```
Run modified simulation scenario
```{r}
ABsim_sealMortFishRun <- rsim.run(ABsim_sealMortFish, method = "RK4", years = 1:25)
```
Plot results
```{r}
rsim.plot(ABsim_sealMortFishRun)
# Plot all species
rsim.plot(ABsim_sealMortFishRun)
# Plot fish
rsim.plot(ABsim_sealMortFishRun, spname = c("cod", "whiting", "mackerel", "anchovy", "shrimp"))
# Plot only seals and cod
rsim.plot(ABsim_sealMortFishRun, spname = c("seals", "cod"))
```


## Seals starting in year 2 and sealers harvest cut in half five years later (year 7) but seals are a strong top-down predator (3x the predation pressure of the base case)
```{r}
ABsim_sealMort <- adjust.forcing(ABsim, parameter = "ForcedMort", group = 'seals', sim.year = 2:25, value = 1.5)
ABsim_sealMortTopDown <- adjust.forcing(ABsim_sealMort, parameter = 'ForcedPrey', group = 'seals', sim.year = 1:25, value = 3) # top-down predator specified for entire scenario
ABsim_sealMortTopDownFish <- adjust.fishing(ABsim_sealMortTopDown, parameter = "ForcedEffort", group = 'sealers', sim.year = 7:25, value = 0.5)
```
Run modified simulation scenario
```{r}
ABsim_sealMortTopDownFishRun <- rsim.run(ABsim_sealMortTopDownFish, method = "RK4", years = 1:25)
```
Plot results
```{r}
# Plot all species
rsim.plot(ABsim_sealMortTopDownFishRun)
# Plot fish
rsim.plot(ABsim_sealMortTopDownFishRun, spname = c("cod", "whiting", "mackerel", "anchovy", "shrimp"))
# Plot only seals and cod
rsim.plot(ABsim_sealMortTopDownFishRun, spname = c("seals", "cod"))
```


## Seals starting in year 2 and sealers harvest is NOT ADJUSTED seals are a strong top-down predator (3x the predation pressure of the base case)
```{r}
ABsim_sealMort <- adjust.forcing(ABsim, parameter = "ForcedMort", group = 'seals', sim.year = 2:25, value = 1.5)
ABsim_sealMortTopDown <- adjust.forcing(ABsim_sealMort, parameter = 'ForcedPrey', group = 'seals', sim.year = 1:25, value = 3) # top-down predator specified for entire scenario
```
Run modified simulation scenario
```{r}
ABsim_sealMortTopDownRun <- rsim.run(ABsim_sealMortTopDown, method = "RK4", years = 1:25)
```
Plot results
```{r}
# Plot all species
rsim.plot(ABsim_sealMortTopDownRun)
# Plot fish
rsim.plot(ABsim_sealMortTopDownRun, spname = c("cod", "whiting", "mackerel", "anchovy", "shrimp"))
# Plot only seals and cod
rsim.plot(ABsim_sealMortTopDownRun, spname = c("seals", "cod"))
```

# Calculating economic value after each simulation

Assuming that cod sells for \$3/pound, mackerel \$2/pound, whiting \$1/pound, shrimp \$4/pound, and anchovies \$0.50/pound

```{r}

# List of fish
fish_species <- c('cod', 'whiting','mackerel', 'anchovy', 'shrimp')
# List of fish prices per pound
fish_value <- c(3,1,2,0.5,4)

# First get the catches and convert them to pounds
AB.full.catches <- AB.full$annual_Catch[,4:8]*2000

# Number of years in the simulation
years <- 100
# Number of species. Both of these are for the loop
species <- 5


fish_prices <- array(rep(NA,years*species),dim=c(years, species))

for (y in seq(1,years)){
  for (i in seq(1,species)){
    # print(paste(fish_species[i],AB.full.catches[y,i],"value is",AB.full.catches[y,i]*fish_value[i]))
    fish_prices[y,i] <- AB.full.catches[y,i]*fish_value[i]
  }
}

print(paste("Total cod revenue through 100 years: USD",sum(fish_prices[,1])))

total_fish_revenue_over_years <- c()
for (fish in seq(1,5)){
  total_fish_revenue_over_years <- append(total_fish_revenue_over_years,sum(fish_prices[,fish]))
}

fish_df <- data.frame(fish_type=fish_species, fish_value=total_fish_revenue_over_years)

png(here("plots","total_fishing_revenue_over_100_years.png"),units='in', height=5, width=7, res=300)
ggplot(fish_df, aes(x=fish_type, y=fish_value)) + geom_bar(stat="identity") + xlab("Total Revenue (USD per km-2) over 100 years") +
  ylab("Species")
dev.off()


print(paste("Total revenue over 100 years from all 5 species: USD ",sum(fish_prices)))
```